---
title       : R and Google Visualization
subtitle    : Tampere University, 19 August 2015
author      : Daniel Fischer
job         : Natural Resources Institute Finland, Green Technology
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
license     : by-nc-sa
github      :
  user      : fischuu
  repo      : fischuu.github.io/presentations/RAndGoogleVisualization
             
---

## Motivation for googleVis

* Due to R's age, some of its basic functions are not up-to-date anymore

* However, as R is modular, user keep developing new packages that improve the basic functionality

* One very popular R-library is [ggplot2](http://ggplot2.org/), that was presented earlier today

* Google has also a large library for programmers that includes different types of algorithms and graphical presentations

* Programmers can utilize these google libraries via standardized interfaces, so-called API's

* Some R-users have created now R-libraries that make the graphic functions provided by google easily accessible via R

```{r setOptions, echo=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```

--- 

## Overview of googleVis

* [googleVis](http://code.google.com/p/google-motion-charts-with-r/) is a package for [R](http://www.r-project.org/) and provides an interface between R and the [Google Chart Tools](https://developers.google.com/chart/)

* The functions of the package allow users to visualise data with the Google Chart Tools without uploading their data to Google

* The output of googleVis functions is html code that contains the data and references to JavaScript functions hosted by Google

* To view the output a browser with an internet connection is required, the actual chart is rendered in the browser; some charts require Flash

* See also: **Using the Google Visualisation API with R**, 
  [The R Journal, 3(2):40-44, December 2011](http://journal.r-project.org/archive/2011-2/RJournal_2011-2_Gesmann+de~Castillo.pdf) and googleVis [package vignette](http://cran.r-project.org/web/packages/googleVis/vignettes/googleVis.pdf)

---

## Usage of googleVis

* The googleVis package is very user friendly and easy to use

* The most difficult part is to get the data to R (what is, with some practice also an easy task)

* A typical workflow would like this:

```{r workflow, eval=FALSE}
library(googleVis)
myData <- data.frame(time = c("JAN", "FEB", "MAR"), 
                 val1 = c(10, 13, 14), 
                 val2 = c(23, 12, 32))
Line <- gvisLineChart(myData)
plot(Line)

```

--- .segue .dark

## Google Diagram Chart Tools

---

## Line chart example
  
```{r linechart, eval=FALSE}
plot(gvisLineChart(myData))
```

```{r linechart.intern, results='asis', echo=FALSE}
library(googleVis)
myData <- data.frame(time = c("JAN", "FEB", "MAR"), 
                 val1 = c(10, 13, 14), 
                 val2 = c(23, 12, 32))
plot(gvisLineChart(myData, options=list(width=950, height=520)))
```

---

## Column chart example
  
```{r columnchart, eval=FALSE}
plot(gvisColumnChart(myData))
```

```{r columnchart.intern, results='asis', echo=FALSE}
library(googleVis)
myData <- data.frame(time = c("US", "EUR", "ASIA"), 
                 val1 = c(10, 13, 14), 
                 val2 = c(23, 12, 32))
plot(gvisColumnChart(myData, options=list(width=950, height=520)))
```

---

## Candlestick chart example I

* Having stock-market data like this, Candlestick charts help to visualize the dynamic of the corresponding stock

```{r Exports data Candle, comment=NA}
print(OpenClose)
```

---

## Candlestick chart example II
  
```{r candlestickchart, eval=FALSE}
plot(gvisCandlestickChart(OpenClose))
```

```{r candlestickchart.intern, results='asis', echo=FALSE}
library(googleVis)
plot(gvisCandlestickChart(OpenClose, options=list(width=950, height=520)))
```

---


## Gauge chart example I

* Another very easy data set:

```{r Exports data Gauge, comment=NA}
print(CityPopularity)
```

---

## Gauge chart example II
  
```{r gaugechart, eval=FALSE}
plot(gvisGauge(CityPopularity, options=list(min=0, max=800, greenFrom=500,
                               greenTo=800, yellowFrom=300, yellowTo=500,
                               redFrom=0, redTo=300)))
```

```{r gaugechart.intern, results='asis', echo=FALSE}
library(googleVis)
plot(gvisGauge(CityPopularity, options=list(min=0, max=800, greenFrom=500,
                               greenTo=800, yellowFrom=300, yellowTo=500,
                               redFrom=0, redTo=300, width=850, height=420)))

```

--- .segue .dark

## Google Geo Chart Tools

---

## Geo chart example I

* Also plotting maps using googleVis is extremly easy

* The following data is enough to create an intensity map

```{r Exports data, comment=NA}
print(Exports)
```

---

## Geo chart example II

```{r intensitychart, eval=FALSE}
plot(gvisGeoChart(Exports, locationvar="Country", 
                 colorvar="Profit",
                 options=list(projection="kavrayskiy-vii")))
```

```{r intensitychart.intern, results='asis', echo=FALSE}
library(googleVis)
plot(gvisGeoChart(Exports, locationvar="Country", 
                 colorvar="Profit",
                 options=list(projection="kavrayskiy-vii", width=950, height=400)))

```

---

## Geo chart example III - Earthquakes in Oceania

```{r intensitychart2.intern, comment=NA}
data(quakes)
head(quakes)
```

---

## Geo chart example IV - Earthquakes in Oceania

```{r intensitychart3.intern, results='asis', echo=FALSE}
library(googleVis)
require(stats)
data(quakes)
quakes$latlong<-paste(quakes$lat, quakes$long, sep=":")
 
plot(gvisGeoChart(quakes, "latlong", "depth", "mag",
                   options=list(displayMode="Markers", region="009",
                   colorAxis="{colors:['red', 'grey']}",
                   backgroundColor="lightblue", width=950, height=500)))
```

---

## Geo chart example V - Finland

```{r finland.data, results='asis', echo=FALSE}
set.seed(54235)
myData <- data.frame(state.name=c(paste("FI-",1:19,sep="")), Something=sample(1:1000,19))
```

```{r finland.data2, comment=NA}
head(myData)
```


---

## Geo chart example VI - Finland

```{r finland, results='asis', echo=FALSE}
library(googleVis)
myData <- data.frame(state.name=c(paste("FI-",1:19,sep="")), Something=sample(1:1000,19))
GeoStates <- gvisGeoChart(myData, "state.name", "Something",
                          options=list(region="FI", 
                                       displayMode="regions", 
                                       resolution="provinces",
                                       width=900, height=450))
plot(GeoStates)

```

---

## Geo chart example VII - My Running

```{r myRunning.data5, results='asis', echo=FALSE}
myRunning <- read.table("/home//ejo138/git//fischuu.github.io//presentations//RAndGoogleVisualization/runningData.Rda")
```

```{r myRunning.data6, comment=NA}
head(myRunning)
```


---

## Geo chart example VI - My Running

```{r myRunning.data7, echo=TRUE, eval=FALSE}
plot(gvisMap(myRunning, "LatLong" , 
             options=list(mapType='terrain', useMapTypeControl=TRUE)))

```


```{r myRunning, results='asis', echo=FALSE}
library(googleVis)
myRunning <- read.table("/home//ejo138/git//fischuu.github.io//presentations//RAndGoogleVisualization/runningData.Rda", nrows =400)
#myRunning <- Andrew
#RunningMap <- gvisMap(myRunning, "LatLong" , 
#                     options=list(showTip=FALSE, 
#                                  showLine=TRUE, 
#                                  enableScrollWheel=TRUE,
#                                  mapType='terrain', 
#                                  useMapTypeControl=TRUE,
#                                  width=950, height=450))
RunningMap <- gvisMap(myRunning, "LatLong" , 
                     options=list(mapType='terrain', useMapTypeControl=TRUE, width=950, height=450))

plot(RunningMap)

```

---

## Motion chart example
  
```{r motionchartcodeWorldBank, results='asis', echo=FALSE}
library(googleVis)
worldBankData <- read.table("/home//ejo138/git//fischuu.github.io//presentations//RAndGoogleVisualization/worldBankData.Rda")
plot(gvisMotionChart(worldBankData, idvar="country.name", timevar="year", options=list(width=950, height=520)))
```
