---
title       : R and Google Visualization
subtitle    : Tampere University, 19 August 2015
author      : Daniel Fischer
job         : Natural Resources Institute Finland, Green Technology
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
license     : by-nc-sa
github      :
  user      : fischuu
  repo      : RAndGoogleVisualization
---

## Disclaimer

1. I am an autodidact 
2. What I present here works for me
3. Read and follow the official [Google Chart API documentation](https://developers.google.com/chart/) and [Terms of Service](https://developers.google.com/readme/terms)
4. Sometimes you have re-load this presentation for the charts and all slides to appear

---

## Agenda

* Introduction and motivation
* Google Chart Tools
* R package googleVis
  * Concepts of googleVis
  * Case studies

```
{r setOptions, message=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```
                  
--- 

## Hans Rosling: No more boring data

<iframe width="420" height="315" src="http://www.youtube.com/embed/hVimVzgtD6w" frameborder="0" allowfullscreen></iframe>

---

## Motivation for googleVis

* Inspired by Hans Rosling’s talks we wanted to use interactive data visualisation tools to foster the dialogue between data analysts and others

* We wanted moving bubbles charts as well

* The software behind Hans’ talk was bought by Google and integrated as motion charts into their Visualisation API

* Ideally we wanted to use R, a language we knew

* Hence, we had to create an interface between the Google Chart Tools and R

--- 

## Overview of googleVis

* [googleVis](http://code.google.com/p/google-motion-charts-with-r/) is a package for [R](http://www.r-poject.org/) and provides an interface between R and the [Google Chart Tools](https://developers.google.com/chart/)

* The functions of the package allow users to visualise data with the Google Chart Tools without uploading their data to Google

* The output of googleVis functions is html code that contains the data and references to JavaScript functions hosted by Google

* To view the output a browser with an internet connection is required, the actual chart is rendered in the browser; some charts require Flash

* See also: **Using the Google Visualisation API with R**, 
  [The R Journal, 3(2):40-44, December 2011](http://journal.r-project.org/archive/2011-2/RJournal_2011-2_Gesmann+de~Castillo.pdf) and googleVis [package vignette](http://cran.r-project.org/web/packages/googleVis/vignettes/googleVis.pdf)

---

## Introduction to Google Chart Tools

* Google Chart Tools provide a way to visualize data on web sites

* The API makes it easy to create interactive charts

* It uses JavaScript and DataTable / JSON as input

* Output is either HTML5/SVG or Flash

* Browser with internet connection required to display chart

* Please read the Google [Terms of Service](https://developers.google.com/terms/) before you start

---

## Structure of Google Charts

The chart code has five generic parts:

1. References to Google’s AJAX and Visualisation API
2. Data to visualise as a DataTable
3. Instance call to create the chart
4. Method call to draw the chart including options
5. HTML &lt;div&gt; element to add the chart to the page

---

## How hard can it be?

* Transform data into JSON object 
* Wrap some HTML and JavaScript around it 
* Thus, googleVis started life in August 2010


----

## Motion chart example

```{r testchartcode, results='asis'}
plot(1,1)
```

---

## Motion chart example

```{r motionchartcode, results='asis'}
library(googleVis)
options(gvis.plot.tag='chart')
plot(gvisMotionChart(Fruits, "Fruit", "Year", options=list(width=800, height=400)), tag="chart")
```
